---
title: "p8105_hw3_bp2678"
author: "Brady Pham"
date: 2023-10-12
output: github_document
---
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

#QUESTION 2 
```{r}
library(p8105.datasets)
data("brfss_smart2010")

```

```{r}
brfss_smart2010_df = brfss_smart2010 |>
  janitor::clean_names() |>
  rename(States = locationabbr,
          County = locationdesc) |>
  filter(response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) |>
  filter(topic == "Overall Health") |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"),
                      ordered = TRUE)) |>
  arrange(response)

```

I renamed the variables like locationabbr to "States" and locationdesc to "County". I also
organized the responses from poor to excellent where poor starts first. Addtionally, I filter
the responses in topic where it is "overall health".

```{r}
newdataset_2002 <- 
  brfss_smart2010_df |>  
  filter(year == "2002")

answer_2002 <- newdataset_2002 |> 
  count(States) |>  
  filter(n >= 7)
answer_2002
```

There are 36 states that were visted more than 7 times in 2002

```{r}
newdataset_2010 <- 
  brfss_smart2010_df |>  
  filter(year == "2010")

answer_2010 <- newdataset_2010 |> 
  count(States) |>  
  filter(n >= 7)
answer_2010

```

There are 45 states that were visted more than 7 times in 2010


Making the Graph 
```{r}
Excellent_only <- brfss_smart2010_df |>
  filter(response == "Excellent") |>
  group_by(year, States) %>%
  summarize(avg_data_value = mean(data_value, na.rm = TRUE)) |>
  ungroup()

ggplot(Excellent_only, aes(x = year, y = avg_data_value, group = States, color = States)) +
  geom_line() +
  labs(title = "Average Data Value Over Time by State",
       x = "Year",
       y = "Average Data Value") +
  theme_minimal()
```

Made NewYork graph
```{r}
NewYork_only <- brfss_smart2010_df |>
  filter(States == "NY", year %in% c(2006, 2010), response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent"))

ggplot(NewYork_only, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year, scales = "free") +
  labs(title = "Distribution of Data Value in NY State by Response",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
```


#Question 3 

```{r}
nhanes_accel_df = 
  read_csv("DATA/nhanes_accel.csv",) |>
  janitor::clean_names() 
```

```{r}
  nhanes_covar_df = 
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
    sex = recode(
      sex,
      `1` = "male",
      `2` = "female")) |>
  mutate(
  education = recode(
    education,
    `1` = "less than high school",
    `2` = "high school equivalent",
    `3` = "more than high school"))
```

#merge the data set 

```{r}
accel_covar <- inner_join(nhanes_covar_df, nhanes_accel_df, by = "seqn") |>
  filter(age >= 21) |>
  drop_na()

```

