---
title: "p8105_hw3_bp2678"
author: "Brady Pham"
date: 2023-10-12
output: github_document
---
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

#QUESTION 2 
```{r}
library(p8105.datasets)
data("brfss_smart2010")

```

```{r}
brfss_smart2010_df = brfss_smart2010 |>
  janitor::clean_names() |>
  rename(States = locationabbr,
          County = locationdesc) |>
  filter(response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) |>
  filter(topic == "Overall Health") |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"),
                      ordered = TRUE)) |>
  arrange(response)

```

I renamed the variables like locationabbr to "States" and locationdesc to "County". I also
organized the responses from poor to excellent where poor starts first. Addtionally, I filter
the responses in topic where it is "overall health".

```{r}
newdataset_2002 <- 
  brfss_smart2010_df |>  
  filter(year == "2002")

answer_2002 <- newdataset_2002 |> 
  count(States) |>  
  filter(n >= 7)
answer_2002
```

There are 36 states that were visted more than 7 times in 2002

```{r}
newdataset_2010 <- 
  brfss_smart2010_df |>  
  filter(year == "2010")

answer_2010 <- newdataset_2010 |> 
  count(States) |>  
  filter(n >= 7)
answer_2010

```

There are 45 states that were visted more than 7 times in 2010


Making the Graph 
```{r}
Excellent_only <- brfss_smart2010_df |>
  filter(response == "Excellent") |>
  group_by(year, States) %>%
  summarize(avg_data_value = mean(data_value, na.rm = TRUE)) |>
  ungroup()

ggplot(Excellent_only, aes(x = year, y = avg_data_value, group = States, color = States)) +
  geom_line() +
  labs(title = "Average Data Value Over Time by State",
       x = "Year",
       y = "Average Data Value") +
  theme_minimal()
```

Made NewYork graph
```{r}
NewYork_only <- brfss_smart2010_df |>
  filter(States == "NY", year %in% c(2006, 2010), response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent"))

ggplot(NewYork_only, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year, scales = "free") +
  labs(title = "Distribution of Data Value in NY State by Response",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
```


#Question 3 

```{r}
nhanes_accel_df = 
  read_csv("DATA/nhanes_accel.csv",) |>
  janitor::clean_names() 
```

```{r}
  nhanes_covar_df = 
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
    sex = recode(
      sex,
      `1` = "male",
      `2` = "female")) |>
  mutate(
  education = recode(
    education,
    `1` = "less than high school",
    `2` = "high school equivalent",
    `3` = "more than high school"))
```

#merge the data set 

```{r}
accel_covar <- inner_join(nhanes_covar_df, nhanes_accel_df, by = "seqn") |>
  filter(age >= 21) |>
  drop_na()

```


```{r}
men_women <- accel_covar |>
  group_by(education, sex) |>
  summarize(count = n()) |>
  pivot_wider(names_from = sex, values_from = count)

print(men_women)

```

For high school equivalent there are 23 females and 35 males
For less than high school there are 28 females and 27 males 
For more than high school there are 59 females and 56 males

```{r}
age_distribution_plot <- ggplot(accel_covar, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions by Education and Gender",
       x = "Education Category",
       y = "Age") +
  theme_minimal()

print(age_distribution_plot)

median_age_table <- accel_covar |>
  group_by(education, sex) |>
  summarize(median_age = median(age, na.rm = TRUE))


```


In high school equivlaent category it seems that females are older than males. The 
median age for females is around 62 while for males the median age seems to be 53. 

In less than high school category it  seems to be fairly equal with distribution that males distribution tend to be a little bit more wider than females. The median age for females seems to be around 61 while males is 60. 

In more than high school category it seems to be fairly equal with both gender with males having a little bit wider of a distribution compared to females. The media age for females seems to be 41 while males seems to be 43.5. 


```{r}
accel_covar |>
  mutate(total_actitivty = rowSums(across(c(min1:min1440)))) |>
  ggplot(aes(x = age, y = total_actitivty, group = sex, color = sex)) + 
  geom_point(alpha = 0.5) + 
  facet_wrap(. ~ education) + 
  geom_smooth(se = FALSE) +
  theme_bw()
```


